---
title: Testing Equality of Multiple Expressions
author: Tyler Littlefield
date: '2018-04-03'
slug: testing-equality-of-multiple-expressions
categories: []
tags: []
---

Credit to BrodieG: https://stackoverflow.com/questions/27325005

This post is essentially an extension of a previous post I [wrote](https://tylurp.rbind.io/2018/02/13/measuring-performance/). The only addition is a cool idea for testing near equality of outputs. I'm calling this an "equality matrix", a matrix of methods that displays whether or not they are equal to each other. One use case is benchmarking. As we benchmark multiple solutions to a single problem, testing whether or not the outputs are equal becomes more time consuming as the `all.equal` function only takes two solutions at a time.

Consider the following problem: _extract all numbers in a vector that have a non zero value after the decimal._

We could do this a few ways:

```{r}
# vector
x <- c(0.0, 0.5, 1.000, 1.5, 1.6, 1.7, 1.75, 2.0, 2.4, 2.5, 3.0, 74.0)

# create objects for testing equality of output
integer_method <- x[as.integer(x) != x]
trunc_method <- x[trunc(x) != x]
round_method <- x[round(x) != x]
mod_method <- x[x %% 1 != 0]
floor_method <- x[floor(x) != x]
```

Now instead of testing every combination to make sure the outputs are equal we can create a matrix that tests all possible combinations at once:

```{r, include=FALSE}
options(width = 10000) 
```

```{r}
# create an equality matrix
methods_vec <- c("integer_method", "trunc_method", "round_method", "mod_method", "floor_method")
objs <- mget(methods_vec)
outer(objs, objs, Vectorize(all.equal))
```

That's it! With just a few lines of code we can test the equality of multiple solutions and print the result nicely. For fun, let's benchmark :smile:


If you've read my previous post on benchmarking, this should all be familiar. First we load the necessary libraries and then we create vectors of different size to test how each solution handles small, medium, and large data:

```{r, message=FALSE}
library(ggplot2)        # plotting
library(patchwork)      # plot multiple plots
library(dplyr)          # using this for pipe functionality 
library(tidyr)          # for gather function when creating the plot I have in mind

# create vectors of different sizes
x <- c(0.0, 0.5, 1.000, 1.5, 1.6, 1.7, 1.75, 2.0, 2.4, 2.5, 3.0, 74.0)
xs <- rep(x, 1e2)
xm <- rep(x, 1e3)
xl <- rep(x, 1e4)
```

Now we benchmark:

```{r}
# benchmark small vector
bms_xs <- bench::mark(
  integer_method = xs[as.integer(xs) != xs],
  trunc_method = xs[trunc(xs) != xs],
  round_method = xs[round(xs) != xs],
  mod_method = xs[xs %% 1 != 0],
  floor_method = xs[floor(xs) != xs]
)

# benchmark medium vector
bms_xm <- bench::mark(
  integer_method = xm[as.integer(xm) != xm],
  trunc_method = xm[trunc(xm) != xm],
  round_method = xm[round(xm) != xm],
  mod_method = xm[xm %% 1 != 0],
  floor_method = xm[floor(xm) != xm]
)

# benchmark large vector
bms_xl <- bench::mark(
  integer_method = xl[as.integer(xl) != xl],
  trunc_method = xl[trunc(xl) != xl],
  round_method = xl[round(xl) != xl],
  mod_method = xl[xl %% 1 != 0],
  floor_method = xl[floor(xl) != xl]
)
```

Then create the plot:

```{r, message=FALSE, fig.height=15}
p1 <- autoplot(bms_xs) + 
  labs(title = "Small Vector") +
  theme(text = element_text(family = "Source Code Pro"))

p2 <- autoplot(bms_xm) + 
  labs(title = "Medium Vector") +
  theme(text = element_text(family = "Source Code Pro"))

p3 <- autoplot(bms_xl) + 
  labs(title = "Large Vector") +
  theme(text = element_text(family = "Source Code Pro"))

vectors <- data.frame(
  small = length(xs),
  medium = length(xm),
  large = length(xl)
)

p4 <- vectors %>% 
  gather() %>% 
  ggplot(aes(key, value)) +
  geom_col(color = "black", fill = "white") +
  labs(title = "Vector Lengths", x = "vector", y = "length") +
  theme(text = element_text(family = "Source Code Pro"))

p1 / p2 / p3 
```




